---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r importing_libraries, message=FALSE}

library(ggplot2)
library(dplyr)
library(knitr)

```

## Loading and preprocessing the data

```{r reading_data}

personal.mov <- read.csv("activity.csv")

```

## What is mean total number of steps taken per day?
```{r questions}

num.steps.per.day <- personal.mov %>% group_by(date) %>% summarise(total.no.steps = sum(steps))

ggplot(data=num.steps.per.day, aes(total.no.steps)) + geom_histogram() +
  ggtitle("Histogram - total number of steps per day") +
  xlab("Total number of steps") +
  ylab("Frequency")

print("Mean and Median values:")

summary(num.steps.per.day$total.no.steps)
```

## What is the average daily activity pattern?
```{r}

num.steps.per.interval <- personal.mov %>% group_by(interval) %>% summarise(mean.no.steps = mean(steps, na.rm=TRUE))

ggplot(data=num.steps.per.interval, aes(x=interval,y=mean.no.steps)) + geom_line()

max.interval <- num.steps.per.interval %>% 
                filter(mean.no.steps == max(mean.no.steps)) %>% 
                select(interval)

cat("Interval with max. num. of steps (in average):",max.interval$interval)
```

## Imputing missing values
```{r}

num.nas <- sum(is.na(personal.mov))

get.mean.num.steps.per.interval <- function(steps.per.interval.df, interv) {
  interval.row <- filter(steps.per.interval.df,interval == interv)
  return(as.double(interval.row[1,2]))
}

personal.mov.complete <- personal.mov
personal.mov.complete <- merge(personal.mov,num.steps.per.interval)
personal.mov.complete$steps <- ifelse(is.na(personal.mov.complete$steps),
                            personal.mov.complete$mean.no.steps,
                            personal.mov.complete$steps)

num.steps.per.day2 <- personal.mov.complete %>% group_by(date) %>% summarise(total.no.steps = sum(steps))

ggplot(data=num.steps.per.day2, aes(total.no.steps)) + geom_histogram() +
  ggtitle("Histogram - total number of steps per day (with imputed data)") +
  xlab("Total number of steps") +
  ylab("Frequency")

print("Mean and Median values:")

summary(num.steps.per.day2$total.no.steps)

```

### Do these values differ from the estimates from the first part of the assignment?

As we can see, the mean value is the same, but the median increased in 10 steps.

### What is the impact of imputing missing data on the estimates of the total daily number of steps?

The height of the histogram for daily total number of steps has increased, as can be seen by the change in the y-axis scale change, as expected. Then, the mean was shifted right.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
personal.mov.complete$weekday <- weekdays(as.POSIXct(personal.mov.complete$date))
personal.mov.complete$day.type <- as.factor(ifelse(personal.mov.complete$weekday %in% c("Saturday","Sunday"),"weekend","weekday"))

num.steps.per.day.type <- personal.mov.complete %>% group_by(day.type, interval) %>% summarize(mean.steps = mean(steps))

plot <- ggplot(data=num.steps.per.day.type, aes(x=interval,y=mean.steps, colour=day.type)) + geom_line()

plot + facet_grid(. ~ day.type)

```


